---
title: "American Slavery and Juneteenth"
author: "Joshua de la Bruere"
date: "6/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweaks to ggplot, and custom themes on github use devtools::install_github("delabj/delabj") to install) 
+ wesanderson has some interesting color palettes
+ gt a newish library by the tidyverse team on the grammar of tables


```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(wesanderson)
library(forcats)
library(gt)

blackpast <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/blackpast.csv')
census <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/census.csv')
slave_routes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/slave_routes.csv')
african_names <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/african_names.csv')


```

Today, I feel like it would interesting to practice with gt, though I'm very unfamiliar with the datasets and how much I'll be able to work with the new package. 

First things first, I want to look at the data

```{r examin data}
blackpast %>% glimpse()
blackpast %>% head()


census %>% glimpse()
census %>% summary()


slave_routes %>% glimpse()
slave_routes %>% head()


african_names %>% glimpse()
african_names %>% head()

```

If I do pursue working with gt today, the census data seems poised for that. 