---
title: "Exmen Data!"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweaks to ggplot, and custom themes on github use devtools::install_github("delabj/delabj") to install) 
+ wesanderson has some interesting color palettes

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(wesanderson)
library(forcats)
library(ggridges)



comic_bechdel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/comic_bechdel.csv')

character_visualization <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/character_visualization.csv')

characters <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/characters.csv')

xmen_bechdel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/xmen_bechdel.csv')

covers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/covers.csv')

issue_collaborators <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/issue_collaborators.csv')

locations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-30/locations.csv')

```

I'm particularly excited by this data set cause the X-Men are some of my favorite comic heros. 


A few of these data sets seem to be fairly straight forward based on the description alone. I want to start by looking at the bechdel test datasets. I have a  feeling that they wont be passing in general.


```{r}
comic_bechdel %>%
  drop_na(writer, pass_bechdel) %>%
  ggplot(aes(x=writer, fill= pass_bechdel))+
  geom_bar(position = position_dodge())

comic_bechdel %>%
  drop_na(issue,series, pass_bechdel) %>%
  ggplot(aes(x=issue, fill= pass_bechdel))+
  facet_wrap(~series, scales="free_x")+
  geom_bar(position = position_dodge())



xmen_bechdel %>%
  drop_na(issue, pass_bechdel) %>%
  ggplot(aes(x=issue, fill= pass_bechdel))+
  geom_bar(position = position_dodge())
```
I'm noticing that the xmen data set has a few issues that are duplicated. I'll use {`janitor`} to see the dupes

```{r}
xmen_bechdel %>% 
  janitor::get_dupes()
xmen_bechdel %>% 
  select(issue) %>%
  janitor::get_dupes()

xmen_bechdel %>%
  distinct(issue, .keep_all= TRUE) %>%
  drop_na(issue, pass_bechdel) %>%
  ggplot(aes(x=as.factor(issue), fill= pass_bechdel))+
  geom_bar()+
  theme_delabj_dark(title_font = "Arial",
                    main.text.color = "#FCB76D",
                    sub.text.color = "#FCB76D",
                    axis.text.color = "#FCB76D", 
                    title.size = 18)+
  delabj::legend_none()+
  labs(title = "<b style = 'color:#FCB76D'> The Uncanny X-MEN</b> ",
       subtitle = "<span style ='color:#FCB76D'> Comic Issues <b style='color:#008F9B'>Passing</b> or <b = style='color:#AC78BA'> Failing </b> the Bechdel test",
       caption= "Data: {claremontrun}\nViz:@delabjl"
       )+
  scale_fill_manual(values = c("#AC78BA","#008F9B"))+
  delabj:::gridlines_off()+
  theme(plot.title = element_markdown(),
        plot.subtitle = element_markdown())+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y =element_blank(),
        axis.ticks.y=element_blank())+
  theme(plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black", colour = NA, size=0),
        panel.border = element_blank())+
  theme(plot.title.position = "panel")+
  theme(plot.caption = element_text(size=8, 
                                    family = "Poppins", 
                                    color = "#FCB76D"))

  ggsave("xmen_bechdel.png", height = 5, width = 10, dpi = 320,device = "png",  type = "cairo")

```

I may have spent a little more time and effort than I meant to on this plot. I haven't even gotten to any of the other data!

```{r}
covers %>%
  glimpse()

covers %>% 
  group_by(cover_artist) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(10) %>%
  ggplot(aes(y=fct_reorder(cover_artist, n), x=n, fill = n)) +
  geom_bar(stat="identity")+
  labs(
    title="Style of The Uncanny X-MEN",
    subtitle = "The 10 most prolific cover artists", 
    caption= "Data: {claremontrun}\nViz:@delabjl",
    fill= "Number of Covers", 
    y=NULL,
    x= NULL
  )+

  scale_fill_gradient(low="#E64E8D", high = "#EE9537" )+
  theme_delabj_dark(title_font = "Arial",
                    main.text.color = "#FCB76D",
                    sub.text.color = "#FCB76D",
                    axis.text.color = "#FCB76D", 
                    title.size = 18)+
    guides(fill= guide_colorbar(
    title.position = "top",
    barwidth = 40,
    barheight = .5,
    ticks.colour = "black",
    hjust = .5
  ))+
  delabj:::gridlines_minor_off()+
  theme(plot.title = element_markdown(),
        plot.subtitle = element_markdown())+
  theme(plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black", colour = NA, size=0),
        panel.border = element_blank())+
  theme(plot.caption = element_text(size=8, 
                                    family = "Poppins", 
                                    color = "#FCB76D"))+
  theme(axis.text.y = element_text(hjust = 0))
  
  ggsave("artist_top10.png", height = 5, width = 10, dpi = 320,device = "png",  type = "cairo")

```


The character dataframe is super rich as well! (Might come back to do some modeling with this at a later date. )

```{r}
characters %>% glimpse()

characters %>%
  group_by(character) %>%
  summarise(captured = sum (captured))%>%
  ggplot(aes(y=fct_reorder(character, captured), x=captured))+
  geom_bar(stat = "identity")
```

I think this is probably prime for pca or clustering. 

in general, I want to split the character column into alias and real name. 


To clean up characters I'd start with this: 
```{r}
characters %>%
  separate(character,  c("alias","secret_id"), sep = "=")

```
To make it so character names are easier to separate out. A similar sort of approach is needed for the covers data


```{r}
tidy_by_char_covers <- covers %>%
  transmute(issue, cover_artist,
    character = strsplit( characters_visualized,"\r\n")) %>%
  unnest(character) %>%
  mutate(
    character = trimws(character),
    #character = str_replace(character, "Dr.", "Doctor"),
    character = case_when(
    str_detect(character, "Ant-man")                  ~ "Ant-Man",
    str_detect(character, "Colossus")                 ~ "Colossus",
    str_detect(character, "Collosus")                 ~ "Colossus",
    str_detect(character, "Cyclops")                  ~ "Cyclops",
    str_detect(character, "Scott Summers")            ~ "Cyclops",
    str_detect(character, "Dazzler")                  ~ "Dazzler",
    str_detect(character, "Sentinels")                ~ "Sentinels",
    str_detect(character, "Nightcrawler")             ~ "Nightcrawler",
    str_detect(character, "Nightcralwer")             ~ "Nightcrawler",
    str_detect(character, "Nightcrewler")             ~ "Nightcrawler",
    str_detect(character, "Wolverine")                ~ "Wolverine",
    str_detect(character, "Wolverime")                ~ "Wolverine",
    str_detect(character, "Storm")                    ~ "Storm",
    str_detect(character, "Storn")                    ~ "Storm",
    str_detect(character, "Captain Marvel")           ~ "Captain Marvel",
    str_detect(character, "Carol Danvers")            ~ "Captain Marvel",
    str_detect(character, "Professor X")              ~ "Professor X",
    str_detect(character, "Profesor X")               ~ "Professor X",
    str_detect(character, "Proffessor X")             ~ "Professor X",
    str_detect(character, "Mystique")                 ~ "Mystique",
    str_detect(character, "Havok")                    ~ "Havok",
    str_detect(character, "Alex Summers")             ~ "Havok",
    str_detect(character, "Kitty Pryde")              ~ "Kitty Pryde",
    str_detect(character, "Kittty Pryde")             ~ "Kitty Pryde",
    str_detect(character, "Emma Frost")               ~ "Emma Frost",
    str_detect(character, "Hellfire Club S")          ~ "Hellfire Club Soldier",
    str_detect(character, "Hellfire Club s")          ~ "Hellfire Club Soldier",
    str_detect(character, "Corsair")                  ~ "Corsair",
    str_detect(character, "Dr.Doom")                  ~ "Doctor Doom",
    str_detect(character, "Jean Grey")                ~ "Jean Grey",
    str_detect(character, "Dark Phoenix")             ~ "Jean Grey",
    str_detect(character, "Phoenix")                  ~ "Jean Grey",
    str_detect(character, "Lockeed")                  ~ "Lockheed the Dragon",
    str_detect(character, "The Dragon")               ~ "Lockheed the Dragon",
    str_detect(character, "Madelyn Prior")            ~ "Madelyne Pryor",
    str_detect(character, "Madelyn Pryoe")            ~ "Madelyne Pryor",
    str_detect(character, "Magistates")               ~ "Magistates",
    str_detect(character, "Sabertooth")               ~ "Sabertooth",
    str_detect(character, "Sabretooth")               ~ "Sabretooth",
    str_detect(character, "Sebastain Shaw")           ~ "Sebastian Shaw",
    str_detect(character, "Sebastin Shaw")            ~ "Sebastian Shaw",
    str_detect(character, "Sebastian Shaw")           ~ "Sebastian Shaw",
    str_detect(character, "Sentinal")                 ~ "Sentinal",
    str_detect(character, "Sleazoid")                 ~ "Sleazoid",
    str_detect(character, "Starjammers")              ~ "Starjammers",
    str_detect(character, "StarJammers(all of them)") ~ "Starjammers",
    str_detect(character, "Archangel")                ~ "Angel",
    str_detect(character, "Archngel")                 ~ "Angel",
    TRUE ~ character
  )) 


tidy_by_char_covers %>%
  group_by(character) %>%
  count () %>%
  arrange(desc(n)) %>%
  head(10) %>%
  mutate(start = 0) %>%
  ggplot()+
  geom_segment(aes(x=start, xend =n , y =fct_reorder(character, n), yend =fct_reorder(character, n)), color = "#1D90F3")+
  geom_point(aes(x=n, y = character), size =5, color = "#FFFE03")+
  labs(
    title = "The Uncany Faces of The X-MEN",
    subtitle = "The most prominant characters by cover apperences",
    caption= "Data: {claremontrun}\nViz:@delabjl",
    x=NULL, 
    y=NULL
  )+
  theme_delabj_dark(
    title_font = "Arial",
    main.text.color = "#FCB76D",
    sub.text.color = "#FCB76D",
    axis.text.color = "#FCB76D", 
    title.size = 18
  )+
  delabj:::gridlines_minor_off()+
  theme(     axis.title.y=element_blank(),
        axis.ticks.y=element_blank())+
  theme(plot.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "black", colour = NA, size=0),
        panel.border = element_blank())+
  theme(plot.caption = element_text(size=8, 
                                    family = "Poppins", 
                                    color = "#FCB76D"))+
  theme(axis.text.y = element_text(hjust = 0))

  

  ggsave("cover_characters_top10.png", height = 5, width = 10, dpi = 320,device = "png",  type = "cairo")

  
```


