---
title: "Beach Volleyball"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweeks to ggplot, and custom themes)

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(forcats)



vb_matches <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-19/vb_matches.csv', guess_max = 76000)

```

## Data Cleaning

I want to look at how individual players preformed. Normally I'd use pivot longer, but I'm not sure how to do that with multiple columns efficently. 

```{r}
winners_p1 <- vb_matches %>%
  select(
    circuit,
    tournament,
    "tournament_country" = country,
    year,
    date,
    gender,
    match_num,
    "player" = w_player1,
    "birthday" = w_p1_birthdate,
    "age"= w_p1_age, 
    "height" = w_p1_hgt,
    "country" = w_p1_country,
    "total_attacks" = w_p1_tot_attacks,
    "total_kills" = w_p1_tot_kills,
    "total_errors" = w_p1_tot_errors,
    "total_hit_pct" = w_p1_tot_hitpct,
    "total_aces" = w_p1_tot_aces,
    "total_serve_errors" = w_p1_tot_serve_errors,
    "total_blocks" = w_p1_tot_blocks,
    "total_digs" = w_p1_tot_digs
     
    ) %>%
   mutate(status =  "Win", 
          player_num = 1) 


winners_p2 <- vb_matches %>%
  select(
    circuit,
    tournament,
    "tournament_country" = country,
    year,
    date,
    gender,
    match_num,
    "player" = w_player2,
    "birthday" = w_p2_birthdate,
    "age"= w_p2_age, 
    "height" = w_p2_hgt,
    "country" = w_p2_country,
    "total_attacks" = w_p2_tot_attacks,
    "total_kills" = w_p2_tot_kills,
    "total_errors" = w_p2_tot_errors,
    "total_hit_pct" = w_p2_tot_hitpct,
    "total_aces" = w_p2_tot_aces,
    "total_serve_errors" = w_p2_tot_serve_errors,
    "total_blocks" = w_p2_tot_blocks,
    "total_digs" = w_p2_tot_digs
     
    ) %>%
   mutate(status =  "Win", 
          player_num = 1) 


loser_p1 <- vb_matches %>%
  select(
    circuit,
    tournament,
    "tournament_country" = country,
    year,
    date,
    gender,
    match_num,
    "player" = l_player1,
    "birthday" = l_p1_birthdate,
    "age"= l_p1_age, 
    "height" = l_p1_hgt,
    "country" = l_p1_country,
    "total_attacks" = l_p1_tot_attacks,
    "total_kills" = l_p1_tot_kills,
    "total_errors" = l_p1_tot_errors,
    "total_hit_pct" = l_p1_tot_hitpct,
    "total_aces" = l_p1_tot_aces,
    "total_serve_errors" = l_p1_tot_serve_errors,
    "total_blocks" = l_p1_tot_blocks,
    "total_digs" = l_p1_tot_digs
     
    ) %>%
   mutate(status =  "Loss", 
          player_num = 2) 


loser_p2 <- vb_matches %>%
  select(
    circuit,
    tournament,
    "tournament_country" = country,
    year,
    date,
    gender,
    match_num,
    "player" = l_player2,
    "birthday" = l_p2_birthdate,
    "age"= l_p2_age, 
    "height" = l_p2_hgt,
    "country" = l_p2_country,
    "total_attacks" = l_p2_tot_attacks,
    "total_kills" = l_p2_tot_kills,
    "total_errors" = l_p2_tot_errors,
    "total_hit_pct" = l_p2_tot_hitpct,
    "total_aces" = l_p2_tot_aces,
    "total_serve_errors" = l_p2_tot_serve_errors,
    "total_blocks" = l_p2_tot_blocks,
    "total_digs" = l_p2_tot_digs
     
    ) %>%
   mutate(status =  "Loss", 
          player_num = 2) 


player_data <- rbind(winners_p1, winners_p2 %>% rbind(loser_p1%>% rbind(loser_p2))) 


# player_data <- player_data %>% 
#   replace_na(
#     list(
#       total_attacks = 0,
#       total_kills = 0,
#       total_errors = 0,
#       total_hit_pct = 0,
#       total_aces = 0,
#       total_serve_errors = 0,
#       total_blocks = 0,
#       total_digs = 0
#     )
#     
#   )

```

## EDA

```{r}
player_data %>%
  summary

```

players by country seemed like an interesting place to start, but there was mostly just a lot of US players

```{r}

player_data %>% 
  group_by(country) %>%
  count() %>%
  ungroup() %>%
  mutate(country=fct_reorder(country, n)) %>%
  arrange(desc(n)) %>%
  head() %>%
  ggplot(aes(y=country, x= n))+
  geom_bar(stat="identity")

```


```{r}
player_data %>%
  group_by(status, gender) %>%
  summarize(mean_attacks=mean(total_attacks)) %>%
  ggplot(aes(y=status, x= mean_attacks))+
  geom_bar(aes(fill=gender), stat="identity", position = "dodge")

player_data %>%
  group_by(status, gender) %>%
  summarize(mean_kills=mean(total_kills)) %>%
  ggplot(aes(y=status, x= mean_kills))+
  geom_bar(aes(fill=gender), stat="identity", position = "dodge")

player_data %>%
  group_by(status, gender) %>%
  summarize(mean_hit_pct=mean(total_hit_pct)) %>%
  ggplot(aes(y=status, x= mean_hit_pct))+
  geom_bar(aes(fill=gender), stat="identity", position = "dodge")

player_data %>%
  group_by(status, gender) %>%
  summarize(mean_aces=mean(total_aces)) %>%
  ggplot(aes(y=status, x= mean_aces))+
  geom_bar(aes(fill=gender), stat="identity", position = "dodge")

player_data %>%
  group_by(status, gender) %>%
  summarize(mean_serve_errors=mean(total_serve_errors)) %>%
  ggplot(aes(y=status, x= mean_serve_errors))+
  geom_bar(aes(fill=gender), stat="identity", position = "dodge")

player_data %>%
  group_by(status, gender) %>%
  summarize(mean_blocks=mean(total_blocks)) %>%
  ggplot(aes(y=status, x= mean_blocks))+
  geom_bar(aes(fill=gender), stat="identity", position = "dodge")

```


## Building a Viz

Maybe density ridges would be interesting? 
  - for count data this would look weird.
  - trying bar plot

```{r}
library(ggridges)

player_data %>% 
  pivot_longer(cols = c(13:20), names_to = "type", values_to = "number") %>%
  mutate(type = gsub("total_", "", type)) %>%
  filter(type != "hit_pct") %>%
  ggplot(aes(x=number))+
  facet_wrap(~type, ncol =1, scales="free_y")+
  geom_bar(aes(fill=status, color=status),position = "identity",alpha = .5)+
  theme_delabj()+
  scale_fill_delabj()+
  scale_color_delabj()+
  xlim(0,40)+
  labs(x="", y="", 
       title = "What Makes A Winner?", 
       subtitle = "<b style= 'color:#83B692'>Winners</b> and <b style = 'color:#721121'>losers</b> Attributes",
       caption = "Data: Adam Vagnar\nViz @delabjl")+
  legend_none()+
  theme(plot.subtitle = element_markdown())

  
ggsave("winner_att.png", type = "cairo", dpi=320, height= 10, width = 10)


player_data %>% 
  pivot_longer(cols = c(13:20), names_to = "type", values_to = "number") %>%
  mutate(type = gsub("total_", "", type)) %>%
  group_by(gender, type) %>%
  summarize(value = median(number, na.rm = TRUE)) %>%
  mutate(type=fct_reorder(type, value)) %>%
  ggplot(aes(y=type, x=value))+
  geom_bar(aes(fill=gender), stat = "identity", position = "dodge")+
  theme_delabj()+
  scale_fill_delabj("retro")+
  labs(x="", y="", 
       title = "Does The Playstyle of <b style= 'color:#FCB76D'>Men </b> & <b style = 'color:#008F9B'>Women</b> Differ?" ,
        caption = "Data: Adam Vagnar\nViz @delabjl")+
  theme(plot.title = element_markdown())+
  legend_none()


ggsave("gender_att.png", type = "cairo", dpi=320, height= 10, width = 10)

```
