---
title: "African American Achievements"
output: github_document
---

## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweaks to ggplot, and custom themes on github use
+ wesanderson has some interesting color palettes
devtools::install_github("delabj/delabj") to install) 

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(wesanderson)
library(forcats)
library(ggridges)

firsts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-09/firsts.csv')
science <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-09/science.csv')

```

I want to examine the data before getting started
```{r}
firsts %>% glimpse()
science %>% glimpse()

```

I think the firsts is probably the one I want to work with first. 

The data was scraped from wikipedia, so there's likely going to be some cleaning if I want to use names of people. 

I want the first person, so this is a really ugly way of doing that.

```{r}

names <- firsts %>% 
  group_by(category) %>%
  filter(person != "Gracia Real de Santa Teresa de Mose (later named Fort Mose) in Florida", 
         accomplishment != "First separate African-American church",
         accomplishment != "First African-American U.S. military regiment"
   )%>%
  top_n(n=-1, year) %>%
  select(category, person, year) %>%
    mutate(name = case_when(
      str_detect(person, "Jupiter Hammon")         ~ "Jupiter Hammon",
      str_detect(person, "Wentworth Cheswell")     ~ "Wentworth Cheswell",
      str_detect(person, "James Derham")           ~ "James Derham",
      str_detect(person, "Rev. Lemuel Haynes")     ~ "Rev. Lemuel Haynes",
      str_detect(person, "Macon Allen")            ~ "Macon Allen",
      str_detect(person, "Joseph Jenkins Roberts") ~ "Joseph Jenkins Roberts",
      str_detect(person, "Major Taylor")           ~ "Major Taylor",
      str_detect(person, "Absalom Boston")         ~ "Absalom Boston",
      TRUE ~ person
    ))


first_by_cat <- firsts %>% 
  group_by(year, category) %>%
  count(year) %>%
  arrange(desc(n))


  ggplot()+
    geom_point(data = first_by_cat, aes(x=year, y=fct_rev(category), size=n,color=category),alpha=0.65, pch=16)+
    #geom_text(data= names, aes(x=year, y= category, label=name), nudge_y = 0.15)+
    delabj::theme_delabj()+
    labs(color=NULL,
       size = "Number of Firsts", 
       x="",
       y="", 
       title = "African American Achievements",
       subtitle = "Firsts in each category by year")+
    delabj::scale_color_delabj()+
    delabj::legend_none()
  
  ggsave("bubble_firsts.png", dpi=320, height = 5, width = 10, type="cairo")
  
  ggplot(data = firsts, aes(x=year, y=fct_rev(category), fill=category))+
    geom_density_ridges(rel_min_height = 0.005, alpha = .75)+
    delabj::theme_delabj()+
    labs(color=NULL,
       size = "Number of Firsts", 
       x="",
       y="", 
       title = "African American Achievements",
       subtitle = "Number of firsts in each category by year")+
    delabj::scale_fill_delabj()+
    delabj::legend_none()
  
    ggsave("ridgeline_firsts.png", dpi=320, height = 5, width = 10, type="cairo")

```

I ended up not liking how the name felt, so I dropped those. I don't know which chart I like better tbh, they both don't feel super satisfactory to me. I think this is all I have time for this week, but it might be interesting to look at cumulative growth or something like that. 
