---
title: "Australian Pets"
author: "Joshua de la Bruere"
date: "7/21/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweaks to ggplot, and custom themes on github use devtools::install_github("delabj/delabj") to install) 

```{r loading, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(forcats)
library(vapoRwave)
library(ggCyberPunk)
library(magrittr)


animal_outcomes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-21/animal_outcomes.csv')
animal_complaints <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-21/animal_complaints.csv')
brisbane_complaints <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-21/brisbane_complaints.csv')

```


I will start by doing some basic EDA to figure out if there's any cleaning that needs to be done or not. I want to play with the `%T>% operator to see if I can combine my standard summary/glipse starting point.


```{r eda_p1}

animal_complaints %T>%
  glimpse() %>%
  mutate(across(where(is.character), as.factor)) %>%
  summary()


animal_outcomes %T>%
  glimpse() %>%
  mutate(across(where(is.character), as.factor)) %>%
  summary()

brisbane_complaints %T>%
  glimpse() %>%
  mutate(across(where(is.character), as.factor)) %>%
  summary()


```
How many types of compants are there? What about by animal

```{r eda_p2}
animal_complaints %>%
  group_by(`Animal Type`, `Complaint Type`) %>%
  count() %>%
  knitr::kable()

```


What about outcomes?
```{r eda_p3}
animal_outcomes %>%
  group_by(animal_type, outcome) %>%
  summarise(total = sum(Total)) %>%
  knitr::kable()

animal_outcomes %>%
  group_by(animal_type, outcome) %>%
  summarise(total = sum(Total)) %>%
  ggplot(aes(y=outcome, x=total))+
  facet_wrap(~animal_type)+
  geom_bar(stat = "identity")+
  theme_minimal()

animal_outcomes %>%
  ggplot(aes(x=year, y=Total, fill = outcome))+
  geom_bar(stat = "identity", position = position_dodge())+
  theme_minimal()+
  legend_bottom()
  

```

I'm not actually sure if I need to clean anything outright. I think I'll be lumping some categorical variables together to avoid having a lot of options. I will pivot the outcomes longer and maybe replace the values.


```{r data cleaning}

region_names <- tribble(
~region_abb, ~region,
'ACT',  'Australian Capital Territory',
'NSW', 	'New South Wales',
'NT', 	'Northern Territory',
'QLD', 	'Queensland',
'SA', 	'South Australia',
'TAS', 	'Tasmania',
'VIC', 	'Victoria',
'WA', 	'Western Australian'
)


animal_complaints %<>%
  janitor::clean_names()

outcomes_cleaned <- animal_outcomes %>%
  # I'm only interested in Cats and Dogs vs Others
  mutate(animal_type = case_when(
    str_detect(animal_type, "Cat") ~ "cat",
    str_detect(animal_type, "Dog") ~ "dog",
    TRUE ~ "other"
  )) %>% 
  pivot_longer(cols = ACT:WA, names_to = "region_abb", values_to = "number") %>%
  left_join(region_names)
  


```


I kinda want to maybe do a statebins style plot. I recently made a pacakge for Africa countries, so I'm familiar with the code. 

<details>

<summary> Setting Up statebins </summary>


I pretty much just copy pasted my Africa code, replaced all africa with austrailia, and then built the shape.

```{r setup_state_bins}
geom_countrybins_australia <- function(
  mapping = NULL,
  data = NULL,
  border_col = "white",
  border_size = 2,
  lbl_size = 3,
  dark_lbl = "black",
  light_lbl = "white",
  radius = grid::unit(6, "pt"),
  ...,
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE){

  ggplot2::layer(
    data = data,
    mapping = mapping,
    stat = "identity",
    geom = Geomcountrybinsaustralia,
    position = "identity",
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      border_col = border_col,
      border_size = border_size,
      lbl_size = lbl_size,
      dark_lbl = dark_lbl,
      light_lbl = light_lbl,
      radius = radius,
      na.rm = na.rm,
      ...
    )
  )
  }


#' @rdname geom_countrybins_australia
#' @export
Geomcountrybinsaustralia <- ggplot2::ggproto(
  "Geomcountrybinsaustralia", ggplot2::Geom,
  default_aes = ggplot2::aes(
    fill = "grey20",
    colour = NA,
    size = 0.1,
    linetype = 1,
    country = "country",
    label="abrv_3_letter",
    angle = 0,
    hjust = 0.5,
    vjust = 0.5,
    alpha = NA,
    family = "",
    fontface = 1,
    lineheight = 1.2
    ),
  extra_params = c(
    "na.rm",
    "width",
    "height"
    ),
  setup_data = function(
    data,
    params) {
    country_data <- data.frame(data, stringsAsFactors=FALSE)

    if (max(nchar(country_data[,"country"])) <= 2) {
      merge.x <- "abrv_2_letter"
    }
    else if (max(nchar(country_data[,"country"])) <= 3){
      merge.x <- "abrv_3_letter"
    }
    else {
      merge.x <- "country"
    }
    country_data <- validate_countries(country_data, "country", merge.x, ignore_dupes=TRUE)
    ctry.dat <- merge(b_country_coords,
                      country_data,
                      by.x=merge.x,
                      by.y="country",
                      all.y=TRUE,
                      sort=TRUE)

    ctry.dat$width  <- ctry.dat$width  %||% params$width  %||% ggplot2::resolution(ctry.dat$x, FALSE)
    ctry.dat$height <- ctry.dat$height %||% params$height %||% ggplot2::resolution(ctry.dat$y, FALSE)
    transform(ctry.dat,
              xmin = x - width / 2,
              xmax = x + width / 2,
              width = NULL,
              ymin = y - height / 2,
              ymax = y + height / 2,
              height = NULL
              ) -> xdat
    xdat
    },
  required_aes = c("country", "fill"),
  draw_panel = function(self, data,
                        panel_params,
                        coord,
                        border_col = "white",
                        border_size = 2,
                        lbl_size = 3,
                        dark_lbl = "black",
                        light_lbl = "white",
                        radius = grid::unit(6, "pt")) {
    tile_data <- data
    tile_data$colour <- border_col
    tile_data$size <- border_size

    text_data <- data
    text_data$label <- data$abrv_3_letter
    text_data$fill <- NA
    text_data$size <-  lbl_size
    text_data$colour <- .sb_invert(
      data$fill,
      dark_lbl,
      light_lbl
      )
    coord <- coord_equal()
    grid::gList(
      GeomRtile$draw_panel(tile_data, panel_params, coord, radius),
      ggplot2::GeomText$draw_panel(text_data, panel_params, coord)
      ) -> grobs
    ggname("geom_countrybins_australia", grid::grobTree(children = grobs))

    },
  draw_key = ggplot2::draw_key_polygon

)


geom_rrect <- function(mapping = NULL, data = NULL,
                       stat = "identity", position = "identity",
                       radius = grid::unit(6, "pt"),
                       ...,
                       na.rm = FALSE,
                       show.legend = NA,
                       inherit.aes = TRUE) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomRrect,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      radius = radius,
      na.rm = na.rm,
      ...
    )
  )
}

GeomRrect <- ggplot2::ggproto("GeomRrect", ggplot2::Geom,

                              default_aes = ggplot2::aes(
                                colour = NA, fill = "grey35", size = 0.5, linetype = 1, alpha = NA
                              ),

                              required_aes = c("xmin", "xmax", "ymin", "ymax"),

                              draw_panel = function(self, data, panel_params, coord,
                                                    radius = grid::unit(6, "pt")) {

                                coords <- coord$transform(data, panel_params)

                                lapply(1:length(coords$xmin), function(i) {

                                  grid::roundrectGrob(
                                    coords$xmin[i], coords$ymax[i],
                                    width = (coords$xmax[i] - coords$xmin[i]),
                                    height = (coords$ymax[i] - coords$ymin)[i],
                                    r = radius,
                                    default.units = "native",
                                    just = c("left", "top"),
                                    gp = grid::gpar(
                                      col = coords$colour[i],
                                      fill = alpha(coords$fill[i], coords$alpha[i]),
                                      lwd = coords$size[i] * .pt,
                                      lty = coords$linetype[i],
                                      lineend = "butt"
                                    )
                                  )

                                }) -> gl

                                grobs <- do.call(grid::gList, gl)

                                ggname("geom_rrect", grid::grobTree(children = grobs))

                              },

                              draw_key = ggplot2::draw_key_polygon

)

geom_rtile <- function(mapping = NULL, data = NULL,
                       stat = "identity", position = "identity",
                       radius = grid::unit(6, "pt"),
                       ...,
                       na.rm = FALSE,
                       show.legend = NA,
                       inherit.aes = TRUE) {
  ggplot2::layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomRtile,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      radius = radius,
      na.rm = na.rm,
      ...
    )
  )
}

GeomRtile <- ggplot2::ggproto("GeomRtile", GeomRrect,

                              extra_params = c("na.rm", "width", "height"),

                              setup_data = function(data, params) {
                                data$width <- data$width %||% params$width %||% ggplot2::resolution(data$x, FALSE)
                                data$height <- data$height %||% params$height %||% ggplot2::resolution(data$y, FALSE)

                                transform(data,
                                          xmin = x - width / 2,  xmax = x + width / 2,  width = NULL,
                                          ymin = y - height / 2, ymax = y + height / 2, height = NULL
                                )
                              },

                              default_aes = ggplot2::aes(
                                fill = "grey20", colour = NA, size = 0.1, linetype = 1, alpha = NA
                              ),

                              required_aes = c("x", "y"),

                              draw_key = ggplot2::draw_key_polygon

)
ggname <- function(prefix, grob) {
  grob$name <- grid::grobName(grob, prefix)
  grob
}


.sb_invert <- function(hex_color, dark_color="black", light_color="white",
                       na_color="white") {

  hex_color <- gsub("#", "", hex_color)

  R <- suppressWarnings(as.integer(paste("0x", substr(hex_color,1,2), sep="")))
  G <- suppressWarnings(as.integer(paste("0x", substr(hex_color,3,4), sep="")))
  B <- suppressWarnings(as.integer(paste("0x", substr(hex_color,5,6), sep="")))

  YIQ <- ((R*299) + (G*587) + (B*114)) / 1000

  return(
    ifelse(is.na(YIQ), na_color,
           ifelse(
             YIQ >= 128, dark_color, light_color)
    )
  )
}

# sanity checks for country values
validate_countries <- function(country_data, country_col, merge.x, ignore_dupes=FALSE) {

  good_country<- country_data[,country_col] %in% country_coords[,merge.x]
  if (any(!good_country)) {
    invalid <- country_data[,country_col][which(!good_country)]
    country_data <- country_data[which(good_country),]
    warning("Found invalid country values: ", invalid)
  }

  if (!ignore_dupes) {
    dupes <- duplicated(country_data[,country_col])
    if (any(dupes)) {
      country_data <- country_data[which(!dupes),]
      warning("Removing duplicate country rows")
    }
  }

  return(country_data)

}

"%||%" <- function(a, b) { if (!is.null(a)) a else b }

.pt <- 2.84527559055118


country_coords <- structure(
  list(
    abrv_2_letter = c(          "NT", "QLD",
                         "WA",   "SA", "NSW", "ACT", 
                                  "VIC",
                                  
                                  "TAS"
                      ),
    abrv_3_letter = c(          "NT", "QLD",
                         "WA",   "SA", "NSW","ACT",
                                       
                                       "VIC", 
                                  
                                  "TAS"
      
                      ),


    country = c(  'Northern Territory', 'Queensland',
                 'Western Australian', 'South Australia','New South Wales','Australian Capital Territory',
                 'Victoria',
                 'Tasmania'
    ),
    col = c(        2L, 3L,
               1,  2L, 3L, 4L,
                        
                        3L, 
                     
                     4L
            ),
    row = c(     1L, 1L,
             2L, 2L, 2L,  2L,
                      
                     3L,
                  5L
    )
  ),
  .Names = c(
    "abrv_2_letter",
             "abrv_3_letter",
             "country",
             "col",
             "row"
    ),
  class = "data.frame", row.names = c(NA, -42L)
)

# Order by Country Name
country_coords <- country_coords[order(country_coords$country),]


# Convert col/row to y and x
b_country_coords <- country_coords
colnames(b_country_coords) <- c(
                                "abrv_2_letter",
                                "abrv_3_letter",
                                "country",
                                "x",
                                "y"
                                )
# Convert Y to negative
b_country_coords$y <- -b_country_coords$y
```
</details>

This is 

```{r}
outcomes_cleaned %>%
  ungroup %>%
  filter(outcome %in% c("Rehomed", "Reclaimed"), year=="2018", 
         animal_type != "other")%>% 
  group_by(region_abb, animal_type) %>%
  summarise(number = sum(number)) %>%
  ungroup %>%
  ggplot(aes(country=region_abb, fill=number))+
    facet_wrap(~animal_type, nrow=1)+
    geom_countrybins_australia()+
  theme_delabj()+
  guides(
    fill = guide_colourbar(
      title.position = "top"
    ))+
  labs(
    title = "With Your Animal It's Home",
    subtitle = "Austrailian States Reclaiming or Rehoming Cats and Dogs in 2018", 
    fill= "Number",
    caption = "Data: RSPCA \n Viz: @delabjl")+
  delabj:::gridlines_off()+
  delabj::legend_left()+
  theme(axis.title = element_blank(), 
        axis.text = element_blank())

ggsave("Homes_Outcome.png", height = 5, width = 10, type="cairo")

outcomes_cleaned %>%
  ungroup %>%
  filter(outcome %in% c("Euthanized"), year=="2018", 
         animal_type != "other")%>% 
  group_by(region_abb, animal_type) %>%
  summarise(number = sum(number)) %>%
  ungroup %>%
  ggplot(aes(country=region_abb, fill=number))+
    facet_wrap(~animal_type, nrow=1)+
    geom_countrybins_australia()+
  theme_delabj()+
  guides(
    fill = guide_colourbar(
      title.position = "top"
    ))+
  scale_fill_gradient(low = "blue", high = "red")+
  delabj::legend_left()+
  labs(
    title = "Euthanized Animals in Australia",
    subtitle = "Austrailian States Euthanizing Cats and Dogs in 2018", 
    fill= "Number",
    caption = "Data: RSPCA \n Viz: @delabjl")+
  delabj:::gridlines_off()

ggsave("Euthanized_Outcome.png", height = 5, width = 10, type="cairo")




```


I'm not thrilled with how this looks, but it'll do for now. I tried animating, but it was a little wonkey and I don't like the aspect ration when it saves. 

I'm curious about outcomes that are "good" vs "bad"


```{r}


order <- outcomes_cleaned %>%
  group_by(region )%>%
  summarise(n=sum(number, na.rm = TRUE)) %>%
  arrange(n) %>% 
  pull(region)



outcomes_cleaned %>%
  mutate(region = fct_rev(factor(region, levels = order))) %>%
  mutate(outcome_type = case_when(
    outcome %in% c( "Euthanized") ~ "Bad", 
    outcome %in% c("Reclaimed", "Rehomed", "Released") ~ "Good",
    TRUE~ "other"
  )) %>%
  group_by(year, region, outcome_type)%>%
  summarise(n= sum(number, na.rm = TRUE)) %>%
  mutate(val = if_else(outcome_type=="Bad", -n, n)) %>%
  filter(outcome_type != "other") %>%
  ggplot(aes(xend=year, x=year, yend=0, y=val, color=outcome_type))+
  facet_wrap(~region, nrow=2)+
  geom_segment()+
  geom_point()+
  delabj::theme_delabj() +
  scale_color_manual(values=c("#721121", "#02075d"))+  
  labs(title="Number Of <span style = 'color:#02075d'> Positive</span>
    and <span style = 'color:#721121'>Negative</span> Outcomes ", 
    subtitle = "For Animials In Australia",
    caption = "Data: RSPCA \n Viz: @delabjl",
    x="", y="")+
  theme(plot.title = element_markdown())

ggsave("outcomes_by_year.png",  height = 5, width = 10, type="cairo")
  
```


