---
title: "Volcanos"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweeks to ggplot, and custom themes)

```{r setup, message=FALSE, warning=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(forcats)

volcano <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/volcano.csv')
eruptions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/eruptions.csv')
events <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/events.csv')
tree_rings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/tree_rings.csv')
sulfur <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-12/sulfur.csv')

```


## EDA

```{r EDA}
library(ggplot2)
volcano %>% summary()
eruptions %>% summary()

world <- map_data("world")


eruptions %>% 
  left_join()
  ggplot(world, aes(x=lat, y=long))+
  geom_sf()
```


I wanted to make a hex grid showing the locations of volcanos (ala Bob Rudis and Pirates!), but I'm stuck on a windows machine, which means I can't update my proj4 library. 
```{r hexgrid}
library(sf)
library(glue)
library(rnaturalearth)




ne_countries(scale = "medium", returnclass = "sf")  -> world

st_read(here::here("data/oceans/ne_110m_ocean.shp"), package="sf") %>%
  select(scalerank, geometry) -> ocean

ocean <- ne_download(scale = 50, type = 'ocean', category = 'physical', returnclass = "sf")

## Reading layer `ne_110m_ocean' from data source `/Users/hrbrmstr/books/30-day-map-challenge/data/ocean/ne_110m_ocean.shp' using driver `ESRI Shapefile'
## Simple feature collection with 2 features and 2 fields
## geometry type:  POLYGON
## dimension:      XY
## bbox:           xmin: -180 ymin: -85.60904 xmax: 180 ymax: 90
## epsg (SRID):    4326
## proj4string:    +proj=longlat +datum=WGS84 +no_defs


if (!file.exists(here::here("data/ocean-grid-250.rds"))) {

  st_make_grid(
    world,
    n = c(250, 250), # grid granularity
    crs = st_crs(world),
    what = "polygons",
    square = FALSE # hex!
  ) -> grid

  grid <- st_sf(index = 1:length(lengths(grid)), grid)

  saveRDS(grid, here::here("data/ocean-grid-250.rds"))

}

grid <- readRDS(here::here("data/ocean-grid-250.rds"))



if (!file.exists(here::here("data/ocean-grid2-250.rds"))) {

  st_make_grid(
    ocean,
    n = c(250, 250), # grid granularity
    crs = st_crs(ocean),
    what = "polygons",
    square = FALSE # hex!
  ) -> grid2

  grid2 <- st_sf(index = 1:length(lengths(grid2)), grid2)

  saveRDS(grid2, here::here("data/ocean-grid2-250.rds"))

}

grid2 <- readRDS(here::here("data/ocean-grid2-250.rds"))

```


I want to get a count of events by volcano. But events is kinda messy. 


```{r cleaning event type}
events_cleaned <- events %>% 
  mutate(event_type=tolower(event_type)) %>%
  mutate(event_type=case_when(grepl("ash", event_type)~ "Ash & Debris",
                              grepl("smoke", event_type)~ "Ash & Debris",
                              grepl("bomb", event_type)~ "Ash & Debris",
                              grepl("lapilli", event_type)~ "Ash & Debris",
                              grepl("tephra", event_type)~ "Ash & Debris",
                              grepl("ejecta", event_type)~ "Ash & Debris",
                              grepl("debris", event_type)~ "Ash & Debris",
                              grepl("earthquake", event_type)~"Earthquake",
                              grepl("seismicity", event_type)~"Earthquake",
                              grepl("tremor", event_type)~"Earthquake",
                              grepl("explosion", event_type)~"Eruption",
                              grepl("eruption", event_type)~"Eruption",
                              grepl("pyroclastic flow", event_type)~"Eruption",
                              grepl("lava", event_type)~"Lava",
                              grepl("degassing", event_type)~"Lava",
                              grepl("formation", event_type)~"Formation",
                              grepl("block", event_type)~"Formation",
                              grepl("fumarolic", event_type)~"Formation",
                              grepl("pumice", event_type)~"Formation",
                              grepl("scoria", event_type)~"Formation",
                              grepl("water", event_type)~"Water Disaster", 
                              grepl("tsunami", event_type)~"Water Disaster",
                              grepl("jokulhaup", event_type)~"Water Disaster",
                              grepl("mud", event_type)~"Water Disaster",
                              grepl("phreatic", event_type)~"Water Disaster",
                              grepl("vei", event_type) ~"VEI", 

                              TRUE~"Other"))
```

```{r events_count}
eruptions %>%
  group_by(volcano_number, latitude, longitude) %>%
  count()->eruptions_count

```








```{r}

ggplot() +
  geom_sf(data = world, fill = "#666666", color = "black", size = 0.0725) +
  geom_hex(data =  eruptions_count, aes(y=latitude, x=longitude),bins=90, alpha=0.8)+
    theme_delabj_dark()+
  scale_fill_viridis_c(option="inferno", 
                        name="", 
                        guide = guide_colorbar(
                          direction = 'horizontal',
                          
                          title.position = 'top', 
                          title.hjust = 0.5, 
                          label.position = "bottom",
                          barwidth = 20,
                          barheight = .5
    
  ))+
  gridlines_off()+
  labs(x="", y="", 
       title="Hot Stuff: Density of Eruptions", 
       caption = "@delabjl")+
  theme(axis.text = element_blank(), 
        plot.caption = element_text(size = 8, hjust = 1), 
        legend.text = element_text(size = 8))


```


If I'm being honest, I think I'm getting stuck on the map. This is far less interesting than I expected it to be. 



Let's try looking at a few other things like activity over time. I'll start with start date. 

```{r activity_over_time}

start_year <- 1800

eruption_only <-  events_cleaned %>% 
  group_by(event_date_year, event_type)%>%
  filter(event_type == "erruption",
         event_date_year >= start_year
         ) %>%
    count() %>%
  ungroup() %>%
  arrange(event_date_year)

all_events<- events_cleaned %>% 
  group_by(event_date_year, event_type)%>%
  filter(event_type != "VEI",
         #event_type != "Other",
         event_date_year >= start_year
         ) %>%
    count() %>%
  ungroup() %>%
  arrange(event_date_year) %>%
 mutate(event_type = factor(event_type, levels = c("Eruption", "Lava", "Ash & Debris", "Earthquake", "Formation", "Water Disaster", "Other") ))


 vol_events_plot <-  ggplot()+
  geom_point(data= all_events, aes(x=event_date_year, y=n, fill = event_type, color=event_type))+
  geom_point(data = eruption_only, aes(x=event_date_year, y=n, fill = event_type, color =event_type))+
  theme_delabj_dark()+
  scale_color_delabj()+
  scale_fill_delabj()+
  guides(col=guide_legend(
    direction = "horizontal",
    title.position = "top",
    label.position = "bottom",
    nrow = 1,
    override.aes = list(shape = 22, size = 5)))+
  labs(
    title="Observed Volcanic Activity by Year", 
    x="", y="", 
  caption =  "Data: The Smithsonian \n Viz: @delabjl"   )+
  theme(legend.title = element_blank())+
  theme(plot.caption = element_text(hjust = 1, size = 10))
 
volcano_cleaned<- volcano %>%
  mutate(last_eruption_year= case_when(
    grepl("Unknown", last_eruption_year)~ NA_character_, 
    TRUE~last_eruption_year)) %>%
  mutate(last_eruption_year = as.numeric(last_eruption_year))
 
vol_loc_plot <- ggplot() +
  geom_sf(data = world, fill = "#666666", color = "black", size = 0.0725)+
  geom_point(data=volcano_cleaned %>% filter(last_eruption_year >= start_year), aes(y=latitude, x=longitude, color = last_eruption_year), alpha = 0.5, size=.5)+
  scale_color_viridis_c(option="magma", 
                        name="", 
                        guide = guide_colorbar(
                          direction = 'horizontal',
                          
                          title.position = 'top', 
                          title.hjust = 0.125, 
                          label.position = "bottom",
                          barwidth = 15,
                          barheight = .5
    
  ))+
  theme_delabj_dark()+
  labs(title="Hot Spots: Locations of Last Eruption Since 1800",
       x="",
       y="")+
  theme(legend.title=element_blank(), 
        plot.caption = element_text(hjust = 1, size = 10))+
  gridlines_off()



ggsave("vol_events.png", plot= vol_events_plot, width = 10, height = 6, dpi=320, type="cairo")
ggsave("vol_loc.png", plot= vol_loc_plot, width = 5, height = 4, dpi=320, type="cairo")


 
```
