---
title: "Coffee Ratings"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweaks to ggplot, and custom themes on github use devtools::install_github("delabj/delabj") to install) 
+ wesanderson has some interesting color palettes

```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(wesanderson)
library(forcats)
library(ggridges)


coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

```


Let's take a look at the data. I see there's  a unit of measurement column. Will this be something I need to convert?

```{r}
summary(coffee_ratings)
glimpse(coffee_ratings)


coffee_ratings %>%
  ggplot(aes(x=unit_of_measurement))+
  geom_bar()

coffee_ratings %>%
  ggplot(aes(y=country_of_origin))+
  geom_bar()

coffee_ratings %>%
  mutate(certification_body = tolower(trimws(certification_body))) %>%
  ggplot(aes(y=certification_body))+
  geom_bar()


coffee_ratings %>%
  mutate(in_country_partner = tolower(trimws(in_country_partner))) %>%
  ggplot(aes(y=in_country_partner))+
  geom_bar()

coffee_ratings %>%
  mutate(in_country_partner = tolower(trimws(in_country_partner))) %>%
  ggplot(aes(y=in_country_partner))+
  geom_bar()


coffee_ratings_cleaned <- coffee_ratings %>%
  mutate(certification_body = tolower(trimws(certification_body))) %>%
  mutate(in_country_partner = tolower(trimws(in_country_partner))) %>%
  mutate(harvest_year = gsub("[^[:digit:].]", "",  harvest_year),
         harvest_year = str_extract(harvest_year, "^.{4}")) %>%
  mutate(country_of_origin = case_when(
    str_detect(country_of_origin, "United States") ~  "United States", 
    str_detect(country_of_origin, "Tanzania") ~  "Tanzania", 
    TRUE ~ country_of_origin
  ))
```
It does look like the altitude has already been converted!

I see a few quality columns, This sort of makes me think about maybe doing a parallel coords plot.

```{r}

df_par <- coffee_ratings_cleaned %>%
   transmute(
     country_of_origin,
        species, 
    processing_method,
     total_cup_points, 
     species, 
     processing_method,
     Aroma = scale(aroma), 
     Acidity = scale(acidity), 
     Aftertaste = scale(aftertaste), 
     Balance = scale(balance), 
     Body = scale(body),
     Flavor = scale(flavor), 
     Moisture = scale(moisture), 
     Sweetness = scale(sweetness),
     Uniformity = scale(uniformity)
         ) %>%
  
  mutate(row_num = row_number()) %>% 
  na.omit() %>%
  pivot_longer(Aroma:Uniformity, names_to = "measure",  values_to = "value") %>%
  filter(measure != "moisture") 


df_par %>%
  ggplot(aes(x=measure, y=value, color = species, group = row_num)) +
  geom_line( size = .5, alpha = .5)+
  theme_minimal()+
  scale_color_delabj()

df_par %>%
  ggplot(aes(x=measure, y=value, color = processing_method, group = row_num)) +
  geom_line( size = .5, alpha = .5)+
  theme_minimal()+
  scale_color_delabj()

```
I'm not particularly liking this. There seems to be a lot of similarities between the species and the processing methods even standardizing the values didn't help much. 

```{r}

coffee_ratings %>% 
  select(
    total_cup_points, 
    species, 
    processing_method,
    aroma:moisture
         ) %>%
  summary()


coffee_ratings %>% 
  select(
    total_cup_points, 
    species, 
    aroma:moisture
         ) %>%
  na.omit() %>%
  ggplot(aes(x= moisture, y=sweetness , color = species))+
  geom_jitter()+
  geom_density2d()


```


It looks like at least from a 2d perspective that there's not much of a clear separation in species except in sweetness, where Robusta makes up a narrow band in the middle. I might just try to look at a few of the attributes using a raincloud plot?



```{r set_up_raincloud}


# Taken from https://micahallen.org/2018/03/15/introducing-raincloud-plots/
"%||%" <- function(a, b) {
  if (!is.null(a)) a else b
}

geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
                             position = "dodge", trim = TRUE, scale = "area",
                             show.legend = NA, inherit.aes = TRUE, ...) {
  layer(
    data = data,
    mapping = mapping,
    stat = stat,
    geom = GeomFlatViolin,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(
      trim = trim,
      scale = scale,
      ...
    )
  )
}

#' @rdname ggplot2-ggproto
#' @format NULL
#' @usage NULL
#' @export
GeomFlatViolin <-
  ggproto("GeomFlatViolin", Geom,
    setup_data = function(data, params) {
      data$width <- data$width %||%
        params$width %||% (resolution(data$x, FALSE) * 0.9)

      # ymin, ymax, xmin, and xmax define the bounding rectangle for each group
      data %>%
        group_by(group) %>%
        mutate(
          ymin = min(y),
          ymax = max(y),
          xmin = x,
          xmax = x + width / 2
        )
    },

    draw_group = function(data, panel_scales, coord) {
      # Find the points for the line to go all the way around
      data <- transform(data,
        xminv = x,
        xmaxv = x + violinwidth * (xmax - x)
      )

      # Make sure it's sorted properly to draw the outline
      newdata <- rbind(
        plyr::arrange(transform(data, x = xminv), y),
        plyr::arrange(transform(data, x = xmaxv), -y)
      )

      # Close the polygon: set first and last point the same
      # Needed for coord_polar and such
      newdata <- rbind(newdata, newdata[1, ])

      ggplot2:::ggname("geom_flat_violin", GeomPolygon$draw_panel(newdata, panel_scales, coord))
    },

    draw_key = draw_key_polygon,

    default_aes = aes(
      weight = 1, colour = "grey20", fill = "white", size = 0.5,
      alpha = NA, linetype = "solid"
    ),

    required_aes = c("x", "y")
  )
```

```{r}


rain_df <-  coffee_ratings_cleaned %>%
   select(
     species, 
     processing_method,
     aroma:moisture
         ) %>%
  mutate(row_num = row_number()) %>% 
  pivot_longer(aroma:moisture, names_to = "measure",  values_to = "value") %>%
  filter(measure != "moisture") 


rain_df %>%
  na.omit() %>%
  filter( measure %!in% c("uniformity", "sweetness", "clean_cup", "cupper_points")) %>%
  ggplot(aes(y = value, x = measure, fill = measure))+
  facet_wrap(~species)+
  geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8, color = "#696969") +
  geom_point(aes( color = measure), position = position_jitter(width = .15), size = .5, alpha = 0.5)+
  geom_boxplot(width = .1, guides = FALSE, outlier.shape = NA, alpha = 0.5, color = "#696969")+
  coord_flip()+
  labs(title = "How do you take it?",
       subtitle = "Distribution of Coffee Profile by species", 
       x=NULL, 
       y=NULL, 
       caption = "Data: Coffee Quality Database\nVis: @delabjl")+
  theme_delabj_dark()+
  legend_none()+
  scale_fill_delabj()+
  scale_color_delabj()+
  theme(strip.background = element_blank(),
        plot.subtitle = element_markdown(), 
        plot.caption = element_text(size = 8))
 

ggsave("coffee_cloud.png", dpi = 320, height = 5, width = 10, type = "cairo")
```


I don't hate the plot, but it's not very interesting imo. I'm still not 100% sure how useful raincloud plots are over violin plots/barplots, or scatter plots. 


What about a profile by country or region? I could look at the profile by the top 10 or so countries to see if there's a difference there. 

```{r}

top_10_countries<- coffee_ratings_cleaned %>% 
  count(country_of_origin, sort = TRUE) %>%
  head(10) %>%
  pull(country_of_origin)

coffee_ratings_cleaned %>%
  filter(country_of_origin %in% top_10_countries
         ) %>%
   transmute(
     country_of_origin,
     total_cup_points, 
     species, 
     processing_method,
     Aroma = scale(aroma), 
     Acidity = scale(acidity), 
     Aftertaste = scale(aftertaste), 
     Balance = scale(balance), 
     Body = scale(body),
     Flavor = scale(flavor), 
     Moisture = scale(moisture), 
     Sweetness = scale(sweetness),
     Uniformity = scale(uniformity)
         ) %>%
  na.omit() %>%
  pivot_longer(Aroma:Uniformity, names_to = "measure",  values_to = "value") %>%
  group_by(country_of_origin, measure) %>%
  summarise( median_val = median(value, na.rm = TRUE)) %>%
  mutate(color = if_else(median_val >=0, "Positive", "negative")) %>%
  ggplot(aes(y=measure, x = median_val, color = color))+
  facet_wrap(~country_of_origin, nrow = 2)+
  geom_segment(aes(x= 0,y=measure, xend = median_val, yend = measure))+
  geom_point()+
  labs(
    title = "Coffee Profiles of The 10 Highest Producing Countries",
    subtitle = "Attributes <b style = 'color: #008F9B'>Above</b> or <b style = 'color:#FCB76D'>Below</b> the worldwide median", 
    y=NULL,
    x=NULL,
    caption = "Data: Coffee Quality Database\nVis: @delabjl"
  )+
  xlim(-1.5,1.5)+
  theme_delabj_dark()+
  scale_color_delabj("retro")+
  theme(strip.background = element_blank(),
        plot.subtitle = element_markdown(), 
        plot.caption = element_text(size = 8))+
  legend_none()
  

ggsave("top_10_profile.png", dpi = 320, height = 5, width = 10, type = "cairo")

```



Hmm. not super happy with any of these plots. 