---
title: "Animal Crossing"
author: "Joshua de la Bruere"
date: "5/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Fetching the data/setup

As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweeks to ggplot, and custom themes)

```{r setup, message=FALSE, warning=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(forcats)

critic <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/critic.tsv')
user_reviews <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/user_reviews.tsv')
items <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/items.csv')
villagers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-05-05/villagers.csv')

```


## EDA

I'm not actually that familiar with animal crossing, I think it's a sims type game? So I might as well take a look at what the data has in store

```{r} 
summary(critic)
summary(user_reviews)
summary(items)
summary(villagers)



critic %>% 
  ggplot(aes(x=grade))+
  geom_histogram(aes(y = stat(density)),  binwidth = 2)+
  geom_density(size = 1)+
  ggtitle("Critic Grades")



user_reviews %>% 
  ggplot(aes(x=grade))+
  geom_histogram(aes(y = stat(density)), binwidth = 1)+
  geom_density(size = 1)+
  ggtitle("User Grades")
  

glimpse(items) %>%
  ggplot(aes(x=buy_value , y= sell_value))+
  geom_point()+
  geom_smooth()


villagers %>%
  ggplot(aes(x=species, fill= gender))+
  geom_bar(position = position_dodge()) +
  facet_wrap(~personality)
  
```
That seems a bit odd, the personalities are split by gender!

```{r}

order <- villagers %>%
  mutate(species = as.factor(species)) %>%
  group_by(species) %>%
  count() %>%
  arrange(desc(n)) %>%
  ungroup()%>%
  mutate(species = as.factor(species)) %>%
  mutate(species =fct_reorder(species, desc(n)))
  


villagers %>%
  mutate(species = factor(species, 
                          levels = order$species), 
         personality = factor(personality, 
                              levels = c("cranky", "jock", "lazy", "smug", 
                                         "normal", "peppy", "snooty", "uchi")
                              )
         ) %>% 
  group_by(species, gender, personality) %>%
  count() %>%
  ggplot(aes(y= species, x=n, fill = gender))+
  geom_bar(stat="identity", position = position_dodge())+
  facet_wrap(~personality, ncol=4)+
  scale_y_discrete(limits = rev(levels(order$species)))+
  labs(title="<b style = 'color:#5D5D5D'>Animal Crossing: Personality & Gender By Species</b>", 
       subtitle = "<i style = 'color:#5D5D5D'> There are <b style = 'color:#7EBBAC'>Male </b> & <b style = 'color:#ED4030'>Female</b> Exclusive Personalities</i>", 
       x="", y="", 
       caption = "<i style = 'color:#5D5D5D'> Data: VillagerDB <br>Viz: @delabjl   </i>")+
  ggcute::theme_sugarpill()+
  theme(plot.title.position = "plot")+
  ggcute::scale_fill_sugarpill(reverse = T)+
  legend_none()+
  theme(plot.subtitle = element_markdown(),
        plot.title = element_markdown(),
        plot.caption = element_markdown())


ggsave(filename = "Personality_gender.png", dpi=320, units = "in", width = 16, height= 9, type ="cairo")

```
