---
title: 'TidyTuesday: College Data'
author: "Joshua de la Bruere"
date: "3/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The first step as always is to load the data form github And to load the libraries we're going to want to use. 
```{r libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(janitor)
library(patchwork)
library(ggtext)
library(delabj)
extrafont::loadfonts(device = "win")
```

```{r read in data, message=FALSE}


tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
```

Now as a general practice I use the `janitor` package to clean my data set names. For this I like to use the `magrittr` packages `%<>%` operator which poth pipes and assigns the result

```{r}
tuition_cost %<>% clean_names()
tuition_income %<>% clean_names()
salary_potential %<>% clean_names()
historical_tuition %<>% clean_names()
diversity_school %<>% clean_names()
```

Let's take a look at the data to see what we're working with. 


```{r}
summary(tuition_cost)
head(tuition_cost)
summary(tuition_income)
head(tuition_income)
summary(salary_potential)
head(salary_potential)
summary(historical_tuition)
head(historical_tuition)
summary(diversity_school)
head(diversity_school)
```

It looks like there could be an interesting difference in growth for mid career vs early career pay. I think it would be interesting to make a plot where you show change in pay and growth using points for starting pay connected by a line representing the change in growth.  We can also use the tuition cost data set to classify them as private or public to see the difference. 

```{r plot pay growth by state}
tmp <- salary_potential %>%
  left_join(tuition_cost) %>%
  mutate(change_in_pay=mid_career_pay-mid_career_pay)%>%
  group_by(state_name, type) %>%
  summarise(early_career_pay = median(early_career_pay),
            mid_career_pay =  median(mid_career_pay), 
            change_in_pay = median(change_in_pay)) %>%
  ungroup()
tmp$state_name <- as.factor(tmp$state_name)
tmp$state_name <- fct_reorder(tmp$state_name, tmp$early_career_pay)
  
p1 <- tmp %>%
pivot_longer(cols = -c(state_name, type), names_to = "timing", values_to = "pay" ) %>%
  
  arrange(desc(timing, pay)) %>%
  filter(timing != "change_in_pay", 
          !is.na(type), 
         type != "For Profit") %>%
  ggplot(aes(y= state_name))+
  geom_line(aes(x=pay))+
  geom_point(aes(x=pay, fill=timing), size = 2, pch = 21)+
  theme_delabj()+
  theme(legend.position = "None")+
  facet_wrap(~type)+
  scale_fill_manual(values= c("#F19953", '#2660A4'))+
  theme(panel.grid.major.y = element_blank())+
  labs(y="", x="Salary", 
       subtitle = "By State",
       title = "Growth From <b style='color:#F19953'>Early Career</b> to <b style=color:'#2660A4'>Mid Career</b> Median Expected Salary", 
       caption = "Data:  Tuitiontracker.org\nViz: @delabj"
       )+
  theme(plot.title = element_markdown(element_markdown(lineheight = 1.2)),
        plot.subtitle = element_markdown(element_markdown(lineheight = 1.1))
        )
ggsave("expected_pay.png",plot=p1, dpi= 300, type = "cairo")  

  
```

```{r plot tution in vs out}



tmp <- tuition_cost %>%
  select(state, type, degree_length, in_state_tuition, out_of_state_tuition) %>%
    filter(
         type == "Public") %>%
  group_by(state, type, degree_length) %>%
  summarise(median_in_state = median(in_state_tuition), 
            median_out_of_state = median(out_of_state_tuition)) %>%
  filter(!is.na(state)) %>%
  mutate(diff= median_out_of_state - median_in_state)

  
tmp$state <- as.factor(tmp$state)
tmp$state <- fct_reorder(tmp$state, tmp$diff)
  
p2 <- tmp %>%
pivot_longer(cols = -c(state, type, degree_length), names_to = "cost_type", values_to = "cost" ) %>%
  filter(cost_type != "diff") %>%
  ggplot(aes(y= state))+
  geom_line(aes(x=cost))+
  geom_point(aes(x=cost, fill=cost_type), size = 2, pch = 21)+
  theme_delabj()+
    theme(legend.position = "None")+
  facet_wrap(~degree_length)+
  scale_fill_manual(values= c("#83B692", '#5B3758' ))+
  theme(panel.grid.major.y = element_blank())+
  labs(y="", x="Tuition", 
       subtitle = "By State",
       title = "Difference between <b style='color:#83B692'>In State Tuition </b> and <b style=color:'#5B3758'>Out of State Tuition </b>", 
       caption = "Data:  Tuitiontracker.org\nViz: @delabj")+
  theme(plot.title = element_markdown(element_markdown(lineheight = 1.2)),
        plot.subtitle = element_markdown(element_markdown(lineheight = 1.1)), 
        
        )+
  theme(plot.margin=margin(6,12,6,12))
ggsave("tuition_by_state.png",plot=p2, dpi= 300, type = "cairo")  

  
```







But does the amount you make from a private school really impact how much you make in general?

```{r}
p3 <- salary_potential %>%
  left_join(tuition_cost) %>%
  group_by(type) %>%
  mutate(mean_tuition = (out_of_state_tuition+ in_state_tuition)/2)%>%
  filter(!is.na(type), 
         type %in% c("Private", "Public"))%>%
  ggplot(aes(group = type, y=mid_career_pay, x=mean_tuition, color=type))+
  geom_point()+
  theme_delabj()+
  scale_color_manual(values= c("#721121", "#525d49"))+
  theme(legend.position = "none")+
  labs(y="", x = "Mean Tuition Cost",
       title= "Does <b style=color:#721121> Private </b> or <b style=color:#525d49> Public </b> Matter?", 
       subtitle = "Expected Mid Career Salary By School Type", 
       caption = "Data:  Tuitiontracker.org\nViz: @delabj")+
    theme(plot.title = element_markdown(element_markdown(lineheight = 1.2)),
        plot.subtitle = element_markdown(element_markdown(lineheight = 1.1)),
        aspect.ratio = 0.5625
        )
  
  ggsave("salary_private_vs_public.png",plot=p3, dpi= 300, type = "cairo")  


```
