---
title: "#TidyTuesday: GDPR Fines"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup
As always I begin by fetching the data from the tidytuesday repo. 

### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)


```{r, message=F}
library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(forcats)
library(delabj)
library(glue)


gdpr_violations <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-21/gdpr_violations.tsv')
gdpr_text <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-21/gdpr_text.tsv')


```



## Examining The Data

Now looking at the data frame `gdpr_text`, I'm not likely to make use of that today, at least for general analysis. Instead I want to gocus on the violations recorded. 

```{r echo=F, message =f}
gdpr_violations
```

There's only 11 variables here and I'm looking at the options and I see there's only a handful of nice data to plot with. There's country name, price, date, controller, and type. 

The article violated looks like it could use some cleaning up, and the country flag is useful, for addding to a plot, but nothing to analyze there. Summary would take some text mining to do anything with that. 

I want to take a look at a summary plot of fine amount by type and country. 
```{r}


gdpr_violations %>%
  group_by(name) %>%
  summarise(total_fines =  sum(price)/1000)%>%
  ggplot(aes(x = total_fines, y = fct_reorder(as.factor(name), total_fines)))+
  geom_bar(stat="identity")+
  theme_delabj()+
  labs(title = "GDPR Fines Collected by Country", 
       y = "",
       x = "Total Fines in 1000 €")


gdpr_violations %>%
  group_by(name) %>%
  count() %>%
  ggplot(aes(x = n, y = fct_reorder(as.factor(name), n)))+
  geom_bar(stat="identity")+
  theme_delabj()+
  labs(title = "Number of Fines Collected by Country", 
       y = "",
       x = "Number of Fines")


gdpr_violations %>%
  group_by(controller) %>%
  summarise(total_fines =  sum(price)/1000)%>%
    arrange(desc(total_fines)) %>%
  head(10) %>%
  ggplot(aes(x = total_fines, y = fct_reorder(as.factor(controller), total_fines)))+
  geom_bar(stat="identity")+
  theme_delabj() +
  labs(title = "GDPR Fines Collected by controller", 
      y="",
       x = "Total Fines in 1000 €")

gdpr_violations %>%
  group_by(type) %>%
  summarise(total_fines =  sum(price)/1000)%>%

  ggplot(aes(x = total_fines, y = fct_reorder(as.factor(type), total_fines)))+
  geom_bar(stat="identity")+
  theme_delabj() +
  labs(title = "GDPR Fines Collected by Type", 
      y="",
       x = "Total Fines in 1000 €")
```


It looks like the type of offence has a few more categories than I expected. I probably should have checked that as a factor before plotting


## Cleaning The Articles

The articles violated seemed like an interesting place to investigate as well as being easier to clean. 

```{r}

gdpr_text %>% group_by(
  article
) %>%
  count()


violated <- gdpr_violations %>%
 filter(str_detect(article_violated, "GDPR"))  %>% # There are 2 unknowns and 2 not classified by gdpr
 mutate(articles_violated = str_split(article_violated, pattern = "\\|"),
         articles_violated = map(articles_violated, 
                                               str_extract, "Art\\.? ?[0-9]+"),
         articles_violated = map(articles_violated, 
                                               ~ sort(unique(.))),
         articles_violated = map(articles_violated, 
                                               str_replace, "Art\\.? ?", "")) %>% #This should get the articles that were violated numerically.
  mutate(num_violated = str_count(articles_violated, ",")+1) #This counts the number of articles violated at once.


```


```{r}
violated %>%
  mutate(date = as.Date(date, format = "%m/%d/%y")) %>%
  arrange(date)%>%
  group_by(date) %>%
  summarise(num_violated = sum(num_violated), 
            price = sum(price)/1000) %>%
  ggplot(aes(x=date, y = num_violated))+
  geom_point(stat="identity",aes(size = price, color = price))+
  theme_delabj()+
  scale_size_continuous(guide = F)+
  scale_x_date(date_breaks = "1 month", limits = NULL)+
  scale_color_delabj("zune", discrete = F)



```
The dates seem a little weird. 


Alright, now lets think about a chart that can explain what we've seen. 


```{r}



violated %>%
  group_by(name) %>%
  summarise(total_fines =  sum(price)/1000,
            num_violations = sum(num_violated))%>%
  arrange(desc(total_fines)) %>%
  head(10)%>%
  ggplot(aes(x = total_fines, y = fct_reorder(as.factor(name), total_fines),fill= num_violations ) )+
  geom_bar(stat="identity")+
  theme_delabj()+
  labs(title = "Who's Collecting the Money?",
       subtitle = "Top 10 Countries Collecting GDPR Fines", 
       y = "",
       x = "Total Fines in 1000 €", 
       caption = "Data: Privacy Affairs\nViz: @delabjl")+
  scale_fill_viridis_c(option = 3, name = "# Violations")+
  legend_right()

  ggsave("Collecting the Money.png", dpi =320, type = "cairo")
  
  
  
  
  gdpr_violations %>%
  group_by(name) %>%
  count() %>%
  ggplot(aes(x = n, y = fct_reorder(as.factor(name), n), fill = n))+
  geom_bar(stat="identity")+
  theme_delabj()+
  labs(title = "Number of Fines Collected by Country", 
        y = "",
       x = "Total Fines in 1000 €", 
       caption = "Data: Privacy Affairs\nViz: @delabjl")+
     scale_fill_viridis_c(option = 3, name = "# Violations")+
  legend_right()

  ggsave("Number of Fines.png", dpi =320, type = "cairo")

```
