---
title: "Broadway SHows"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup
As always I begin by fetching the data from the tidytuesday repo. 
### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweeks to ggplot, and custom themes)

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(forcats)

grosses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/grosses.csv', guess_max = 47525)
synopses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/synopses.csv')
cpi <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/cpi.csv')
pre_1985_starts <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-28/pre-1985-starts.csv')

### TRouble Shooting
grosses_orig <- readr::read_csv('https://raw.githubusercontent.com/tacookson/data/master/broadway-grosses/grosses.csv')

```


I don't have a ton of domain knowledge, so I'm not quite sure what the best option is. Maybe a bumpchart or bar chart race?


```{r standing room only}


grosses %>%
  mutate(week_ending = as.Date(week_ending))%>%
  mutate(month = lubridate::month(week_ending),
          year = lubridate::year(week_ending)) %>%
  filter(year >= 1995) %>%
  mutate(weekly_potential_seats = seats_in_theatre*performances,
         sold_vs_potential = seats_sold/weekly_potential_seats) %>%
  group_by(show) %>%
  summarise(seats_sold = mean(seats_sold), 
            potential_sold = mean(weekly_potential_seats),
            pct_capacity = max(pct_capacity, na.rm = T)) %>%
  mutate(pct_capacity_calc = potential_sold/pct_capacity)%>%
  filter(potential_sold <= seats_sold)%>%
  mutate(show = fct_reorder(show, seats_sold)) %>% 
  arrange(desc(seats_sold)) %>%
  head(5) %>%
   pivot_longer(cols = c(potential_sold, seats_sold), names_to ="type") %>%
  ggplot(aes(y = show))+
  geom_line(aes(x= value), color = "#AC78BA")+
  geom_point(aes(x= value, color = type), size = 2)+
  theme_delabj()+
  scale_color_manual(values = c( "#83B692","#AC78BA"))+
  labs(title = "Standing Room Only:",
       subtitle = "Top 5 shows that had more <b style = 'color:#AC78BA'> seats sold </b> than 
       <b style = 'color:#83B692'> potential seats</b>",
       x="Seats",
       y="", 
       caption = "Data: Playbill/TidyTuesday\n
      Viz: @delabjl")+
  theme(plot.subtitle = element_markdown())+
  legend_none()

ggsave("standing_room.png", dpi=320, type="cairo")
  

```


```{r Least Filled}


grosses %>%
  mutate(week_ending = as.Date(week_ending))%>%
  mutate(month = lubridate::month(week_ending),
          year = lubridate::year(week_ending)) %>%
  filter(year >= 1995) %>%
  mutate(weekly_potential_seats = seats_in_theatre*performances,
         sold_vs_potential = seats_sold/weekly_potential_seats) %>%
  group_by(show) %>%
  summarise(seats_sold = mean(seats_sold), 
            potential_sold = mean(weekly_potential_seats),
            pct_capacity = max(pct_capacity, na.rm = T)) %>%
  mutate(pct_capacity_calc = potential_sold/pct_capacity)%>%
  filter(potential_sold >= seats_sold)%>%
    filter(seats_sold > 0 )%>%
  mutate(show = fct_reorder(show, pct_capacity)) %>% 
  arrange(pct_capacity) %>%
  head(5) %>%
   pivot_longer(cols = c(potential_sold, seats_sold), names_to ="type") %>%
  ggplot(aes(y = show))+
  geom_line(aes(x= value), color = "#AC78BA")+
  geom_point(aes(x= value, color = type), size = 2)+
  theme_delabj()+
  scale_color_manual(values = c( "#83B692","#AC78BA"))+
  labs(title = "Private Showings:",
       subtitle = "The 5 Shows with the lowest mean <b style = 'color:#AC78BA'> Seats Sold </b> to 
       <b style = 'color:#83B692'> Potential Seats</b> ratio",
       x="Seats",
       y="", 
       caption = "Data: Playbill/TidyTuesday\n
      Viz: @delabjl")+
  theme(plot.subtitle = element_markdown())+
  legend_none()

ggsave("private.png", dpi=320, type="cairo")
  

```

```{r popular}
grosses %>%
  mutate(week_ending = as.Date(week_ending))%>%
  mutate(month = lubridate::month(week_ending),
          year = lubridate::year(week_ending)) %>%
  filter(year >= 1995) %>%
  mutate(weekly_potential_seats = seats_in_theatre*performances,
         sold_vs_potential = seats_sold/weekly_potential_seats) %>%
  group_by(show) %>%
  summarise(seats_sold = mean(seats_sold), 
            potential_sold = mean(weekly_potential_seats),
            pct_capacity = max(pct_capacity, na.rm = T)) %>%
  mutate(pct_capacity_calc = potential_sold/pct_capacity)%>%
    filter(seats_sold > 0 )%>%
  mutate(show = fct_reorder(show, desc(pct_capacity))) %>% 
  arrange(desc(pct_capacity) )%>%
  head(5) %>%
   pivot_longer(cols = c(potential_sold, seats_sold), names_to ="type") %>%
  ggplot(aes(y = show))+
  geom_line(aes(x= value), color = "#AC78BA")+
  geom_point(aes(x= value, color = type), size = 2)+
  theme_delabj()+
  scale_color_manual(values = c( "#83B692","#AC78BA"))+
  labs(title = "Popular:",
       subtitle = "The 5 Shows with the highest mean <b style = 'color:#AC78BA'> Seats Sold </b> to 
       <b style = 'color:#83B692'> Potential Seats</b> ratio",
       x="Seats",
       y="", 
       caption = "Data: Playbill/TidyTuesday\n
      Viz: @delabjl")+
  theme(plot.subtitle = element_markdown())+
  legend_none()

ggsave("popular.png", dpi=320, type="cairo")

```
