---
title: "BBC Rap Poll"
author: "Joshua de la Bruere"
date: "4/14/2020"
output:  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fetching the data/setup
As always I begin by fetching the data from the tidytuesday repo. Considering the topic today, I am also considering using {spotifyR} to get some more data on the songs. 

### Packages:
These are the package I plan on using today:

+ tidyverse (It's what this is all about)
+ spotifyr (Extra data from spotify's API)
+ ggtext (A wonderful package by [@ClausWilke](https://twitter.com/ClausWilke) that enables better control over text rendering)
+ patchwork (By the amazing [@thomasp85](https://twitter.com/thomasp85) allows for easy combining of plots)
+ janitor (makes it easy to clean the names of a data set.)
+ forcats (easy work with factors)
+ delabj (A personal package that includes a few tweeks to ggplot, and custom themes)

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
library(spotifyr)
library(ggtext)
library(patchwork)
library(janitor)
library(delabj)
library(forcats)
library(ggchicklet)

polls <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv')
rankings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/rankings.csv')


```

let's take a look at the data

```{r}
polls
rankings
```


I see that the way this poll was calculated, used a multi choise method, which means there could be an interesting exploration of different counting methods. I want to take a look to see if this is a viable option. 

```{r}
polls %>%
  arrange(critic_name)

```

I'm pretty sure that something can be done with  borda count and first past the post. I don't have all pairwise comparisions, so I can't do condorcet comparisions without making a few assumptions. 

```{r bbcMethod}
bc <- rankings %>%
  arrange(desc(points)) %>%
  head(10) %>%
  mutate(title = as.factor(title)) %>%
  mutate(title =  fct_reorder(title, points)) %>%
  mutate(winner = if_else(points == max(rankings$points), "Winner", "Loser")) %>%
  ggplot(aes(x=points, y=title, fill = winner))+
  geom_bar(stat="identity")+
  labs(y="", x="Points", 
       title = "BBC Counting Method (Borda Count)", 
       caption = "Data: BBC Music\n Viz: @delabjl")+
  theme_delabj_dark()+
  scale_fill_delabj()+
  legend_none()

ggsave("bc.png", plot = bc, dpi=320, type = "cairo")
bc
```


How this is calclualted
```{r}
combo_bc <- rankings %>%
  arrange(desc(points)) %>%
  head(10) %>%
  mutate(title = as.factor(title)) %>%
  mutate(title =  fct_reorder(title, points)) %>%
  mutate(winner = if_else(points == max(rankings$points), "Winner", "Loser")) %>%
  pivot_longer(cols =  c('n1','n2','n3','n4','n5'),names_to = 'round') %>%
  mutate(points_by_round = case_when(
    round == "n1" ~ value * 10,
    round == "n2" ~ value * 8, 
    round == "n3" ~ value * 6,
    round == "n4" ~ value * 4,
    round == "n5" ~ value * 2,
    TRUE ~ 0
  )) %>%
  ggplot(aes(y=points_by_round, x=title, fill = winner, group = round))+
  geom_chicklet( color  = "#343E48")+
  coord_flip()+
   labs(y="", x="", 
       title = "Contribution By Round With BBC Borda Count", 
       caption = "Data: BBC Music\n Viz: @delabjl")+
  theme_delabj_dark()+
  scale_fill_delabj()+
  legend_none()


supp.labs <- c("Round 1","Round 2", "Round 3", "Round 4", "Round 5")
names(supp.labs) <- c("n1", "n2", "n3", "n4", "n5")


by_round_bc <- rankings %>%
  arrange(desc(points)) %>%
  head(10) %>%
  mutate(title = as.factor(title)) %>%
  mutate(title =  fct_reorder(title, points)) %>%
  mutate(winner = if_else(points == max(rankings$points), "Winner", "Loser")) %>%
  pivot_longer(cols =  c('n1','n2','n3','n4','n5'),names_to = 'round') %>%
  mutate(points_by_round = case_when(
    round == "n1" ~ value * 10,
    round == "n2" ~ value * 8, 
    round == "n3" ~ value * 6,
    round == "n4" ~ value * 4,
    round == "n5" ~ value * 2,
    TRUE ~ 0
  )) %>%
  ggplot(aes(y=points_by_round, x=title, fill = winner, group = round))+
  geom_chicklet( color  = "#343E48")+
  coord_flip()+
   labs(y="Points", x="", 
       title = "Points Gained by Round",
       sub= "Using the Borda Count Method",
       caption = "Data: BBC Music\n Viz: @delabjl")+
  theme_delabj_dark()+
  scale_fill_delabj()+
  legend_none()+
  facet_wrap(~round, ncol = 2, labeller = labeller(round = supp.labs))+
  theme(strip.text = element_text(colour = "#D6D6D6"))+
  theme(plot.caption = element_text(size = 8))

ggsave("combo_borda.png", plot = combo_bc, dpi=320, type = "cairo")
ggsave("by_round_borda.png", plot = by_round_bc, dpi=320, type = "cairo")
combo_bc
by_round_bc

```


Raw Votes
```{r Raw Vote Count}
rc_chart <- rankings %>%
  arrange(desc(points)) %>%
  head(10) %>%
  mutate(title = as.factor(title)) %>%
  mutate(title =  fct_reorder(title, n)) %>%
  mutate(winner = if_else(n == max(n), "Winner", "Loser")) %>%
  ggplot(aes(x=n, y=title, fill = winner))+
  geom_bar(stat="identity")+
  labs(y="", x="Number of Votes All Rounds", 
       title = "Equal Point Ranked Voting", 
       caption = "Data: BBC Music\n Viz: @delabjl")+
  theme_delabj_dark()+
  scale_fill_delabj()+
  legend_none()

ggsave("rc.png", plot = rc_chart, dpi=320, type = "cairo")
rc_chart


```

```{r FPP}
fpp <- polls %>%
  arrange(critic_name) %>%
  filter(rank == 1) %>%
  group_by(title, artist, gender) %>%
  summarise(votes = sum(rank), 
            percent_vote = sum(rank)/107)
fpp_chart <- fpp %>% arrange(desc(votes)) %>%
  head(10) %>%
  ungroup() %>%
  mutate(title = as.factor(title)) %>%
  mutate(title =  fct_reorder(title, percent_vote)) %>%
  mutate(winner = if_else(votes == max(fpp$votes), "Winner", "Loser")) %>%
  ggplot(aes(y=title, x = percent_vote, fill= winner))+
  geom_bar(stat = "identity")+
  labs(y="", x="% Vote", 
       title = "First Past the Post Method",
       caption = "Data: BBC Music\n Viz: @delabjl")+
  scale_x_continuous(labels = scales::percent_format(accuracy = 1))+
  theme_delabj_dark()+
  scale_fill_delabj()+
  legend_none()

ggsave("fpp.png", plot = fpp_chart, dpi=320, type = "cairo")
fpp_chart

``` 



